---
title: "Outcome analysis report"
subtitle: "This report was developed with assistance from Cursor."
author: "Dr Tom Moore - Statistical Scientist"
date: last-modified
format:
  html:
    output-file: 1_targets_report.html
    embed-resources: true
    include-in-header:
      text: |
        <style>
        .fig-squeeze .figure, .fig-squeeze .quarto-float-figure { margin-bottom: 0 !important; }
        .quarto-title-block .subtitle, .subtitle { font-size: 0.85em; opacity: 0.9; font-style: italic; }
        </style>
    df-print: paged
    toc: true
    toc-location: right
    toc-depth: 5
    theme: united
    highlight-style: tango
    fig-width: 5
    fig-height: 3
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| message: false
#| warning: false
# Setup
# Set root directory so targets store is found (analysis/scripts/pipeline/)
root_dir <- getwd()
pipeline_dir <- file.path(root_dir, "analysis", "scripts", "pipeline")
if (!file.exists(file.path(pipeline_dir, "_targets.R"))) {
  pipeline_dir <- file.path(root_dir, "pipeline")  # when run from analysis/scripts/
}
if (file.exists(file.path(pipeline_dir, "_targets.R"))) {
  root_dir <- pipeline_dir
}
knitr::opts_knit$set(root.dir = root_dir)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(targets)
library(patchwork)
if (!requireNamespace("crew", quietly = TRUE)) {
  suppressMessages(
    capture.output(
      install.packages("crew", repos = "https://cloud.r-project.org/", quiet = TRUE),
      type = "output"
    )
  )
}
tar_unscript()
# Project root: from pipeline dir go up 3; from analysis/scripts go up 2; else use getwd()
project_root <- if (basename(getwd()) == "pipeline") {
  dirname(dirname(dirname(getwd())))
} else if (basename(getwd()) == "scripts" && basename(dirname(getwd())) == "analysis") {
  dirname(dirname(getwd()))
} else {
  getwd()
}
# Plots, tables, and models (always under analysis/output per project rules)
plot_dir <- file.path(project_root, "analysis", "output", "plots")
tables_dir <- file.path(project_root, "analysis", "output", "tables")
models_dir <- file.path(project_root, "analysis", "output", "models")
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(tables_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(models_dir, recursive = TRUE, showWarnings = FALSE)
# Activate renv so this session and crew workers see the project library (fixes missing janitor/tidybayes/ggdag in tar_make)
if (requireNamespace("renv", quietly = TRUE) && file.exists(file.path(project_root, "renv/activate.R"))) {
  renv::activate(project_root)
}
# CmdStan: on Linux (cluster) use known path if it exists; else use CMDSTAN_PATH if set; otherwise default
cluster_cmdstan <- "/powerplant/workspace/hrltxm/workbench-k8s/stan/mod_stan/cmdstan-2.36.0"
if (Sys.info()["sysname"] == "Linux" && dir.exists(cluster_cmdstan) && !nzchar(Sys.getenv("CMDSTAN_PATH"))) {
  Sys.setenv(CMDSTAN_PATH = cluster_cmdstan)
}
if (nzchar(Sys.getenv("CMDSTAN_PATH")) && requireNamespace("cmdstanr", quietly = TRUE)) {
  cmdstanr::set_cmdstan_path(Sys.getenv("CMDSTAN_PATH"))
}
```
```{targets example-globals, tar_globals = TRUE}
# Globals
tar_option_set(format = "rds", 
               controller = crew::crew_controller_local(name = "HAL 9000", workers = 5, seconds_idle = 10),
               memory = "transient",
               garbage_collection = TRUE,
               seed = 42,
               packages = c(
                 "tidyverse", "janitor", "modelr", "tidybayes", "brms",
                 "bayesplot", "ggdag", "cmdstanr", "targets", "patchwork", "testthat"))
```


```{targets raw_data_file}
#| include: false
#| echo: false
# Path and existence of airquality.csv in analysis/data/import/. When exists is FALSE, raw_data simulates.
# Encoding existence in the target value ensures downstream targets (e.g. raw_data) use the right branch.
# If you add or remove airquality.csv in import/, run targets::tar_invalidate(raw_data_file) then tar_make() to refresh.
tar_target(raw_data_file, {
  pr <- if (basename(getwd()) == "pipeline") dirname(dirname(dirname(getwd()))) else getwd()
  if (basename(getwd()) == "scripts" && basename(dirname(getwd())) == "analysis") pr <- dirname(dirname(getwd()))
  path <- file.path(pr, "analysis", "data", "import", "airquality.csv")
  list(path = path, exists = file.exists(path))
})
```

```{targets data_source}
#| include: false
#| echo: false
# Data source mode: use_import when airquality.csv exists, use_anonymised when pipeline simulates (e.g. from structure.rds).
tar_target(data_source, {
  list(use_import = raw_data_file$exists, use_anonymised = !raw_data_file$exists)
})
```

```{targets simulate_structure}
#| include: false
#| echo: false
# Structure for simulation: read from analysis/data/simulate/structure.rds.
# Accepts (1) a list with n and columns (type/min/max per column), or (2) a data frame (derived from str(df)-style info: nrow, column types, ranges).
tar_target(simulate_structure, {
  pr <- if (basename(getwd()) == "pipeline") dirname(dirname(dirname(getwd()))) else getwd()
  if (basename(getwd()) == "scripts" && basename(dirname(getwd())) == "analysis") pr <- dirname(dirname(getwd()))
  sim_dir <- file.path(pr, "analysis", "data", "simulate")
  struct_file <- file.path(sim_dir, "structure.rds")
  default <- list(
    n = 153L,
    columns = list(
      v1 = list(type = "numeric", min = 1, max = 168),
      v2 = list(type = "numeric", min = 7, max = 334),
      v3 = list(type = "numeric", min = 2.3, max = 20.7),
      v4 = list(type = "numeric", min = 56, max = 97),
      v5 = list(type = "integer", min = 5L, max = 9L),
      v6 = list(type = "integer", min = 1L, max = 31L)
    )
  )
  if (!file.exists(struct_file)) return(default)
  x <- readRDS(struct_file)
  if (is.data.frame(x)) {
    n <- nrow(x)
    columns <- lapply(x, function(col) {
      if (is.factor(col)) {
        list(type = "factor", levels = as.character(levels(col)))
      } else if (is.integer(col)) {
        r <- range(col, na.rm = TRUE)
        list(type = "integer", min = r[1L], max = r[2L])
      } else if (is.numeric(col)) {
        r <- range(col, na.rm = TRUE)
        list(type = "numeric", min = r[1L], max = r[2L])
      } else {
        list(type = "numeric", min = 0, max = 1)
      }
    })
    names(columns) <- names(x)
    return(list(n = n, columns = columns))
  }
  if (is.list(x) && !is.null(x$columns)) {
    return(list(n = x$n, columns = x$columns))
  }
  default
})
```

```{targets raw-data}
#| include: false
#| echo: false
# Raw data: read CSV (rename to v1..v6 using mapping in analysis/data) or simulate from structure (v1..v6)
tar_target(raw_data, {
  path <- raw_data_file$path
  if (isTRUE(raw_data_file$exists)) {
    d <- readr::read_csv(path, show_col_types = FALSE)
    pr <- dirname(dirname(dirname(dirname(path))))
    mapping_path <- file.path(pr, "analysis", "data", "import", "column_mapping.R")
    if (!file.exists(mapping_path)) stop("column_mapping.R not found in analysis/data/import/")
    map_env <- new.env()
    source(mapping_path, local = map_env)
    orig_to_v <- get("orig_to_v", envir = map_env)
    nms <- names(d)
    for (i in seq_along(orig_to_v)) {
      idx <- match(names(orig_to_v)[i], nms)
      if (!is.na(idx)) nms[idx] <- unname(orig_to_v[i])
    }
    names(d) <- nms
    return(d)
  }
  struct <- simulate_structure
  n <- if (!is.null(struct$n)) as.integer(struct$n) else 100L
  cols <- struct$columns
  if (is.null(cols) || length(cols) == 0) {
    stop("No structure to simulate from; add analysis/data/simulate/structure.rds or provide columns in default.")
  }
  out <- lapply(names(cols), function(name) {
    spec <- cols[[name]]
    type <- if (!is.null(spec$type)) spec$type else "numeric"
    if (identical(type, "factor")) {
      levs <- if (!is.null(spec$levels)) spec$levels else character(0)
      if (length(levs) == 0) levs <- NA_character_
      sample(levs, size = n, replace = TRUE)
    } else if (identical(type, "integer")) {
      min_val <- if (!is.null(spec$min)) spec$min else 1L
      max_val <- if (!is.null(spec$max)) spec$max else 31L
      sample(seq(min_val, max_val), size = n, replace = TRUE)
    } else {
      min_val <- if (!is.null(spec$min)) spec$min else 0
      max_val <- if (!is.null(spec$max)) spec$max else 1
      runif(n, min = min_val, max = max_val)
    }
  })
  names(out) <- names(cols)
  tibble::as_tibble(out)
}, packages = "readr")
```
```{targets wrangling}
#| include: false
#| echo: false
# Wrangling: log-transform v1; keep complete cases on v1, v2
tar_target(data, {
  raw_data %>%
    filter(!is.na(v1), !is.na(v2)) %>%
    mutate(log_v1 = log(v1))
})
```
```{targets data_struct}
#| include: false
#| echo: false
# Structure
tar_target(data_struct, data)
```
```{targets data_summary}
#| include: false
#| echo: false
# Summary
tar_target(data_summary, data %>% summarise(across(everything(), mean)))
```
```{targets stan_code}
#| include: false
#| echo: false
# Stan code (v4 → v3; log_v1 ~ v3 * v4 + v2)
tar_target(stan_code, {
  bf_v3 <- brms::bf(v3 ~ v4)
  bf_v1 <- brms::bf(log_v1 ~ v3 * v4 + v2)
  brms::stancode(bf_v3 + bf_v1 + brms::set_rescor(FALSE), data = data, family = list(gaussian(), gaussian()))
})
```

<!-- Models -->
```{targets fit}
#| include: false
#| echo: false
tar_target(fit, {
  if (requireNamespace("cmdstanr", quietly = TRUE)) {
    path <- Sys.getenv("CMDSTAN_PATH")
    if (!nzchar(path) && Sys.info()["sysname"] == "Linux") {
      cluster_cmdstan <- "/powerplant/workspace/hrltxm/workbench-k8s/stan/mod_stan/cmdstan-2.36.0"
      if (dir.exists(cluster_cmdstan)) path <- cluster_cmdstan
    }
    if (nzchar(path)) cmdstanr::set_cmdstan_path(path)
  }
  options(mc.cores = parallel::detectCores())
  # Structural equations: (1) v4 → v3; (2) log_v1 ~ v3 * v4 + v2
  bf_v3 <- brms::bf(v3 ~ v4)
  bf_v1 <- brms::bf(log_v1 ~ v3 * v4 + v2)
  mod_formula <- bf_v3 + bf_v1 + brms::set_rescor(FALSE)
  
  priors <- c(brms::prior(normal(0, 1), class = Intercept),
              brms::prior(normal(0, 1), class = b),
              brms::prior(normal(0, 1), class = sigma))
  
  tryCatch({
    brms::brm(mod_formula, family = list(gaussian(), gaussian()),
        data = data,
        threads = brms::threading(threads = NULL, grainsize = 1250, static = FALSE),
        backend = "cmdstanr", silent = 0, refresh = 100,
        chains = 4, cores = 4, warmup = 1000, iter = 2000, seed = 42)
  }, error = function(e) {
    brms::brm(mod_formula, family = list(gaussian(), gaussian()),
        data = data,
        backend = "rstan", silent = 0, refresh = 100,
        chains = 4, cores = 4, warmup = 1000, iter = 2000, seed = 42)
  })
})
```

```{targets fit_saved}
#| include: false
#| echo: false
# Save Bayesian fit to analysis/output/models (per project rules); runs after fit so file is always written
tar_target(fit_saved, {
  pr <- if (basename(getwd()) == "pipeline") dirname(dirname(dirname(getwd()))) else getwd()
  if (basename(getwd()) == "scripts" && basename(dirname(getwd())) == "analysis") pr <- dirname(dirname(getwd()))
  models_dir <- file.path(pr, "analysis", "output", "models")
  dir.create(models_dir, recursive = TRUE, showWarnings = FALSE)
  path <- file.path(models_dir, "bayesian_fit.rds")
  saveRDS(fit, path)
  path
}, format = "file")
```

```{targets dag}
#| include: false
#| echo: false
# DAG: v4 (exposure) → v3 (mediator); log_v1 (outcome) ~ v3, v4, v2 (covariate)
tar_target(dag, {
  basic_dag <- dagify(
    v3 ~ v4,
    log_v1 ~ v3,
    log_v1 ~ v4,
    log_v1 ~ v2,
    exposure = "v4",
    outcome = "log_v1",
    coords = list(
      x = c(v4 = 0, v3 = 1, v2 = 0.5, log_v1 = 2),
      y = c(v4 = 1, v3 = 1.1, v2 = 0.4, log_v1 = 1)
    )
  )
  basic_dag_plot <- basic_dag |>
    ggdag::tidy_dagitty() |>
    dplyr::mutate(
      var_type = dplyr::case_when(
        name == "v4" ~ "Exposure",
        name == "log_v1" ~ "Outcome",
        name == "v3" ~ "Mediator",
        name == "v2" ~ "Covariate"
      ),
      label = dplyr::case_match(
        name,
        "v4" ~ "v4",
        "v3" ~ "v3",
        "log_v1" ~ "log(v1)",
        "v2" ~ "v2"
      )
    )
  ggplot2::ggplot(basic_dag_plot, ggplot2::aes(x = x, y = y, xend = xend, yend = yend)) +
    ggdag::geom_dag_edges() +
    ggdag::geom_dag_point(ggplot2::aes(color = var_type), size = 21) +
    ggdag::geom_dag_text(
      ggplot2::aes(label = label),
      color = "white", fontface = "bold", size = 4
    ) +
    ggplot2::scale_color_manual(
      values = c(Exposure = "#882255", Mediator = "#6699CC", Outcome = "#999933", Covariate = "#CC6677"),
      guide = "none"
    ) +
    ggdag::theme_dag() +
    ggplot2::theme(plot.margin = ggplot2::margin(12, 12, 12, 12))
})
```

```{targets pp_check}
#| include: false
#| echo: false
# Validation
tar_target(pp_check, brms::pp_check(fit, "dens_overlay", resp = "logv1") + theme_classic())
```

<!-- Raw plots -->
```{targets plot_ozone_wind}
#| include: false
#| echo: false
tar_target(plot_ozone_wind, {
  v4_cond <- quantile(data$v4, 0.9)
  pred_grid <- data %>%
    data_grid(v3 = seq_range(v3, n = 200), v4 = v4_cond, v2 = mean(v2)) %>%
    add_epred_draws(fit, ndraws = 100) %>%
    filter(.category == "logv1") %>%
    mutate(.epred = exp(.epred))
  ggplot(pred_grid, aes(x = v3, y = .epred)) +
    theme_classic() +
    stat_lineribbon(.width = .95, alpha = 0.8, fill = "#6699CC", color = "black") +
    geom_point(data = data, aes(x = v3, y = v1), shape = 21, size = 2, inherit.aes = FALSE) +
    labs(
      title = "v1 vs v3",
      subtitle = paste0("v4 held at 90th percentile (", round(v4_cond, 1), ")"),
      y = "v1", x = "v3"
    )
})
```

```{targets plot_ozone_temp}
#| include: false
#| echo: false
tar_target(plot_ozone_temp, {
  pred_grid <- data %>%
    data_grid(v4 = seq_range(v4, n = 200), v3 = mean(v3), v2 = mean(v2)) %>%
    add_epred_draws(fit, ndraws = 100) %>%
    filter(.category == "logv1") %>%
    mutate(.epred = exp(.epred))
  v3_val <- round(mean(data$v3), 1)
  ggplot(pred_grid, aes(x = v4, y = .epred)) +
    theme_classic() +
    stat_lineribbon(.width = .95, alpha = 0.8, fill = "#882255", color = "black") +
    geom_point(data = data, aes(x = v4, y = v1), shape = 21, size = 2, inherit.aes = FALSE) +
    labs(
      title = "v1 vs v4",
      subtitle = paste0("v3 and v2 held at mean (", v3_val, ")"),
      y = "v1", x = "v4"
    )
})
```

```{targets plot_ozone_solar_r}
#| include: false
#| echo: false
tar_target(plot_ozone_solar_r, {
  pred_grid <- data %>%
    data_grid(v2 = seq_range(v2, n = 200), v3 = mean(v3), v4 = mean(v4)) %>%
    add_epred_draws(fit, ndraws = 100) %>%
    filter(.category == "logv1") %>%
    mutate(.epred = exp(.epred))
  ggplot(pred_grid, aes(x = v2, y = .epred)) +
    theme_classic() +
    stat_lineribbon(.width = .95, alpha = 0.8, fill = "#CC6677", color = "black") +
    geom_point(data = data, aes(x = v2, y = v1), shape = 21, size = 2, inherit.aes = FALSE) +
    labs(
      title = "v1 vs v2",
      subtitle = "v3 and v4 held at mean",
      y = "v1", x = "v2"
    )
})
```


<!-- Effect plots -->
```{targets eff_wind_by_temp}
#| include: false
#| echo: false
tar_target(eff_wind_by_temp, {
  cond <- data.frame(v4 = quantile(data$v4, probs = c(0.25, 0.75)), v2 = mean(data$v2))
  v4_vals <- round(cond$v4, 1)
  ce <- conditional_effects(fit, effects = "v3", conditions = cond, re_formula = NA)
  dfs <- ce[sapply(ce, is.data.frame)]
  c_eff <- if (length(dfs) >= 2L) dfs[[2L]] else dfs[[1L]]
  c_eff <- c_eff %>% mutate(across(c(estimate__, lower__, upper__), exp))
  ggplot(c_eff, aes(x = effect1__, y = estimate__, ymin = lower__, ymax = upper__,
                    fill = cond__, color = cond__)) +
    theme_classic() +
    theme(legend.position = "bottom") +
    geom_ribbon(alpha = 0.3, colour = NA) +
    geom_line(linewidth = 1.5) +
    scale_fill_manual(values = c("#882255", "#6699CC"), name = "v4") +
    scale_color_manual(values = c("#882255", "#6699CC"), name = "v4") +
    labs(
      title = "Effect of v3 on v1",
      subtitle = paste0("v4 held at 25th and 75th percentiles (", paste(v4_vals, collapse = " and "), ")"),
      y = "v1", x = "v3"
    )
})
```

```{targets eff_wind_by_solar_r}
#| include: false
#| echo: false
tar_target(eff_wind_by_solar_r, {
  cond <- data.frame(v2 = quantile(data$v2, probs = c(0.25, 0.75)), v4 = mean(data$v4))
  v2_vals <- round(cond$v2, 1)
  ce <- conditional_effects(fit, effects = "v3", conditions = cond, re_formula = NA)
  dfs <- ce[sapply(ce, is.data.frame)]
  c_eff <- if (length(dfs) >= 2L) dfs[[2L]] else dfs[[1L]]
  c_eff <- c_eff %>% mutate(across(c(estimate__, lower__, upper__), exp))
  ggplot(c_eff, aes(x = effect1__, y = estimate__, ymin = lower__, ymax = upper__,
                    fill = cond__, color = cond__)) +
    theme_classic() +
    theme(legend.position = "bottom") +
    geom_ribbon(alpha = 0.3, colour = NA) +
    geom_line(linewidth = 1.5) +
    scale_fill_manual(values = c("#CC6677", "#44AA99"), name = "v2") +
    scale_color_manual(values = c("#CC6677", "#44AA99"), name = "v2") +
    labs(
      title = "Effect of v3 on v1",
      subtitle = paste0("v2 held at 25th and 75th percentiles (", paste(v2_vals, collapse = " and "), ")"),
      y = "v1", x = "v3"
    )
})
```

```{targets unit_tests}
#| include: false
#| echo: false
# Analysis unit tests (testthat): data and model checks; returns tibble with test name and passed/failed
tar_target(unit_tests, {
  library(testthat)
  library(tibble)
  run_one <- function(name, expr) {
    tryCatch(
      { force(expr); tibble(test = name, passed = TRUE) },
      error = function(e) tibble(test = name, passed = FALSE)
    )
  }
  dplyr::bind_rows(
    run_one("Data: no NA in v1", testthat::expect_true(all(!is.na(data$v1)))),
    run_one("Data: required columns (v1, v2, v3, v4, log_v1)", testthat::expect_true(all(c("v1", "v2", "v3", "v4", "log_v1") %in% names(data)))),
    run_one("Data: n > 0", testthat::expect_true(nrow(data) > 0L)),
    run_one("Data: v1 positive", testthat::expect_true(all(data$v1 > 0, na.rm = TRUE))),
    run_one("Data: log_v1 is numeric", testthat::expect_true(is.numeric(data$log_v1))),
    run_one("Model: converged (Rhat < 1.05)", testthat::expect_true(all(brms::rhat(fit) < 1.05, na.rm = TRUE))),
    run_one("Model: has v3 and v4 terms", testthat::expect_true(any(grepl("v3", rownames(brms::fixef(fit)))) && any(grepl("v4", rownames(brms::fixef(fit)))))),
    run_one("Model: has v2 term", testthat::expect_true(any(grepl("v2", rownames(brms::fixef(fit)))))),
    run_one("Model: bivariate (2 responses)", testthat::expect_true(length(fit$formula) >= 2L))
  ) %>%
    dplyr::arrange(passed) %>%
    dplyr::mutate(
      passed_display = dplyr::if_else(passed, "\u2191", "\u2717"),
      passed_html = dplyr::if_else(passed, '<span style="color:#16a34a">\u2191</span>', '<span style="color:#dc2626">\u2717</span>')
    )
}, packages = c("testthat", "tidyverse", "brms"))
```

```{targets qmd_file}
#| include: false
#| echo: false
# Track 1_targets.qmd so setup_unit_tests invalidates when the report is edited (e.g. title change)
tar_target(qmd_file, {
  project_root <- if (basename(getwd()) == "pipeline") dirname(dirname(dirname(getwd()))) else getwd()
  if (basename(getwd()) == "scripts" && basename(dirname(getwd())) == "analysis") project_root <- dirname(dirname(getwd()))
  file.path(project_root, "analysis", "scripts", "1_targets.qmd")
}, format = "file")
```

```{targets readme_file}
#| include: false
#| echo: false
# Track README.md so setup_unit_tests invalidates when README is edited
tar_target(readme_file, {
  project_root <- if (basename(getwd()) == "pipeline") dirname(dirname(dirname(getwd()))) else getwd()
  if (basename(getwd()) == "scripts" && basename(dirname(getwd())) == "analysis") project_root <- dirname(dirname(getwd()))
  file.path(project_root, "README.md")
}, format = "file")
```

```{targets project_rules_file}
#| include: false
#| echo: false
# Track project rules so setup_unit_tests invalidates when rules are edited
tar_target(project_rules_file, {
  project_root <- if (basename(getwd()) == "pipeline") dirname(dirname(dirname(getwd()))) else getwd()
  if (basename(getwd()) == "scripts" && basename(dirname(getwd())) == "analysis") project_root <- dirname(dirname(getwd()))
  file.path(project_root, ".cursor", "rules", "project-rules.mdc")
}, format = "file")
```

```{targets install_packages_file}
#| include: false
#| echo: false
# Track install_packages.R so setup_unit_tests invalidates when package list changes
tar_target(install_packages_file, {
  project_root <- if (basename(getwd()) == "pipeline") dirname(dirname(dirname(getwd()))) else getwd()
  if (basename(getwd()) == "scripts" && basename(dirname(getwd())) == "analysis") project_root <- dirname(dirname(getwd()))
  file.path(project_root, "analysis", "scripts", "setup", "install_packages.R")
}, format = "file")
```

```{targets setup_unit_tests}
#| include: false
#| echo: false
# Setup unit tests: compliance with project rules (structure, Cursor, Quarto, git, config)
tar_target(setup_unit_tests, {
  library(testthat)
  library(tibble)
  run_one <- function(name, expr) {
    tryCatch(
      { force(expr); tibble(test = name, passed = TRUE) },
      error = function(e) tibble(test = name, passed = FALSE)
    )
  }
  pr <- if (basename(getwd()) == "pipeline") dirname(dirname(dirname(getwd()))) else getwd()
  if (basename(getwd()) == "scripts" && basename(dirname(getwd())) == "analysis") pr <- dirname(dirname(getwd()))
  qmd_path <- qmd_file
  readme_path <- readme_file
  project_rules_path <- project_rules_file
  install_packages_path <- install_packages_file
  p <- function(...) file.path(pr, ...)
  read_txt <- function(path) paste(readLines(path, warn = FALSE), collapse = " ")

  dirs <- setNames(c("analysis", "analysis/data", "analysis/data/import", "analysis/output", "analysis/output/models", "analysis/output/plots", "analysis/output/tables", "analysis/scripts", "analysis/scripts/pipeline", "analysis/scripts/setup", "docs", ".cursor/rules"),
    c("Directory: analysis/", "Directory: analysis/data/", "Directory: analysis/data/import/", "Directory: analysis/output/", "Directory: analysis/output/models/", "Directory: analysis/output/plots/", "Directory: analysis/output/tables/", "Directory: analysis/scripts/", "Directory: analysis/scripts/pipeline/", "Directory: analysis/scripts/setup/", "Directory: docs/", "Directory: .cursor/rules"))
  dir_tests <- lapply(seq_along(dirs), function(i) run_one(paste0(names(dirs)[i], " exists"), testthat::expect_true(dir.exists(p(dirs[i])))))

  files <- setNames(c(".cursorignore", ".gitignore", "README.md"), c("File: .cursorignore", "File: .gitignore", "File: README.md"))
  file_tests <- lapply(seq_along(files), function(i) run_one(paste0(names(files)[i], " at root"), testthat::expect_true(file.exists(p(files[i])))))
  rproj_test <- list(run_one("File: .Rproj at root", testthat::expect_true(length(list.files(pr, pattern = "\\.Rproj$")) >= 1L)))

  qmd <- if (file.exists(qmd_path)) read_txt(qmd_path) else ""
  git <- function(...) tryCatch(system2("git", c(...), stdout = TRUE, stderr = FALSE), error = function(e) character(0))
  gi <- if (file.exists(p(".gitignore"))) read_txt(p(".gitignore")) else ""
  ci <- if (file.exists(p(".cursorignore"))) read_txt(p(".cursorignore")) else ""

  report_tests <- list(
    run_one("Cursor transparency: subtitle mentions Cursor", testthat::expect_true(grepl("subtitle:.*[Cc]ursor", qmd))),
    run_one("Quarto: title not default", testthat::expect_false(grepl('title:\\s*["\']New Document["\']', qmd))),
    run_one("Quarto: .qmd in analysis/scripts/", testthat::expect_true(length(list.files(p("analysis", "scripts"), pattern = "\\.qmd$")) >= 1L)),
    run_one("Quarto only: no .Rmd in analysis/scripts/", testthat::expect_true(length(list.files(p("analysis", "scripts"), pattern = "\\.Rmd$")) == 0L))
  )
  git_tests <- list(
    run_one("Git: is repository", testthat::expect_true(dir.exists(p(".git")) || identical(git("rev-parse", "--is-inside-work-tree"), "true"))),
    run_one("Git: branch is main", testthat::expect_true(git("branch", "--show-current") %in% c("main", "master"))),
    run_one("Git: remote origin", testthat::expect_true("origin" %in% git("remote", "show")))
  )
  config_tests <- list(
    run_one("README: mentions pipeline/run", testthat::expect_true(grepl("pipeline|quarto|run|quick start", read_txt(readme_path), ignore.case = TRUE))),
    run_one("Project rules: project-rules.mdc", {
      testthat::expect_true(file.exists(project_rules_path))
      testthat::expect_true(grepl("Quarto|analysis|Cursor", read_txt(project_rules_path)))
    }),
    run_one("Output writable: analysis/output/plots", {
      tf <- p("analysis", "output", "plots", ".proj_test"); dir.create(dirname(tf), recursive = TRUE, showWarnings = FALSE)
      writeLines("x", tf); on.exit(unlink(tf, force = TRUE), add = TRUE)
      testthat::expect_true(file.exists(tf)); unlink(tf, force = TRUE)
    }),
    run_one("No top-level scripts/ (per rules)", testthat::expect_false(dir.exists(p("scripts")))),
    run_one(".cursorignore: excludes _targets or output", testthat::expect_true(grepl("_targets|output", ci))),
    run_one("Output writable: analysis/output/tables", {
      tf <- p("analysis", "output", "tables", ".proj_test"); dir.create(dirname(tf), recursive = TRUE, showWarnings = FALSE)
      writeLines("x", tf); on.exit(unlink(tf, force = TRUE), add = TRUE)
      testthat::expect_true(file.exists(tf)); unlink(tf, force = TRUE)
    }),
    run_one("Report YAML: has output-file", testthat::expect_true(grepl("output-file|output_file", qmd))),
    run_one("R version >= 4.0", testthat::expect_true(getRversion() >= "4.0.0"))
  )

  # Pipeline packages (must match tar_option_set(packages = ...) in globals)
  pipeline_packages <- c(
    "tidyverse", "janitor", "modelr", "tidybayes", "brms",
    "bayesplot", "ggdag", "cmdstanr", "targets", "patchwork", "testthat"
  )
  package_tests <- lapply(pipeline_packages, function(pkg) {
    run_one(paste0("Package: ", pkg), {
      ok <- suppressPackageStartupMessages(suppressWarnings(require(pkg, character.only = TRUE, quietly = TRUE)))
      testthat::expect_true(ok)
    })
  })

  software_tests <- list(
    run_one("Software: R available", testthat::expect_true(nzchar(R.version.string))),
    run_one("Software: Quarto (quarto --version)", {
      code <- tryCatch(system2("quarto", "--version", stdout = FALSE, stderr = FALSE), error = function(e) 1L)
      testthat::expect_true(code == 0L)
    })
  )

  out <- dplyr::bind_rows(c(dir_tests, file_tests, rproj_test, report_tests, git_tests, config_tests, package_tests, software_tests))
  out %>% dplyr::arrange(passed) %>%
    dplyr::mutate(
      passed_display = dplyr::if_else(passed, "\u2191", "\u2717"),
      passed_html = dplyr::if_else(passed, '<span style="color:#16a34a">\u2191</span>', '<span style="color:#dc2626">\u2717</span>')
    )
}, packages = c("testthat", "tidyverse"))
```

```{targets report_unit_tests}
#| include: false
#| echo: false
# Report unit tests: setup + analysis summary, Git release readiness, and report/output checks
tar_target(report_unit_tests, {
  pt_setup <- setup_unit_tests
  pt_analysis <- unit_tests
  pr <- if (basename(getwd()) == "pipeline") dirname(dirname(dirname(getwd()))) else getwd()
  if (basename(getwd()) == "scripts" && basename(dirname(getwd())) == "analysis") pr <- dirname(dirname(getwd()))
  git_ok <- pt_setup %>% dplyr::filter(grepl("^Git:", .data$test)) %>% dplyr::pull(.data$passed) %>% all()
  report_path <- file.path(pr, "analysis", "scripts", "1_targets_report.html")
  plot_dir <- file.path(pr, "analysis", "output", "plots")
  models_dir <- file.path(pr, "analysis", "output", "models")
  report_exists <- file.exists(report_path)
  key_plots_ok <- all(file.exists(file.path(plot_dir, c(
    "effect_v3_by_v4.png", "effect_v3_by_v2.png", "causal_dag.png",
    "pp_check.png", "v1_vs_v3_raw.png", "v1_vs_v4_raw.png", "v1_vs_v2_raw.png"
  ))))
  model_saved <- file.exists(file.path(models_dir, "bayesian_fit.rds"))
  dplyr::bind_rows(
    dplyr::tibble(test = "All setup unit tests pass", passed = all(pt_setup$passed)),
    dplyr::tibble(test = "All analysis unit tests pass", passed = all(pt_analysis$passed)),
    dplyr::tibble(test = "Git: release readiness (branch main, clean tree, origin)", passed = git_ok),
    dplyr::tibble(test = "Report output exists", passed = report_exists),
    dplyr::tibble(test = "Key outputs present (plots and model)", passed = key_plots_ok && model_saved)
  ) %>%
    dplyr::mutate(
      passed_html = dplyr::if_else(passed, '<span style="color:#16a34a">\u2191</span>', '<span style="color:#dc2626">\u2717</span>')
    )
}, packages = "dplyr")
```

```{targets setup_unit_tests_ok}
#| include: false
#| echo: false
# Gate: pipeline stops here unless all setup unit tests pass
tar_target(setup_unit_tests_ok, {
  pt <- setup_unit_tests
  if (!all(pt$passed)) {
    failed <- pt$test[!pt$passed]
    stop("Setup unit tests failed: ", paste(failed, collapse = "; "))
  }
  invisible(TRUE)
})
```

<details><summary>Pipeline</summary>

```{r}
#| label: visnet
#| echo: false
#| fig-height: 6
#| fig-width: 9
#| out-width: 75%
tryCatch({
  library(visNetwork)
  tar_visnetwork(
    targets_only = TRUE,
    label = "description",
    label_width = 28,
    level_separation = 120,
    outdated = TRUE
  ) |>
    visNodes(
      font = list(size = 14, face = "sans"),
      shape = "box",
      color = list(
        background = "#f0f4f8",
        border = "#475569",
        highlight = list(background = "#e2e8f0", border = "#334155")
      ),
      margin = 8,
      shadow = TRUE
    ) |>
    visEdges(
      color = list(color = "#94a3b8", highlight = "#64748b"),
      smooth = list(type = "cubicBezier", roundness = 0.4),
      width = 1.5,
      arrows = list(to = list(enabled = TRUE, scaleFactor = 0.6))
    ) |>
    visOptions(
      highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
      nodesIdSelection = FALSE
    ) |>
    visLegend(enabled = FALSE, useGroups = FALSE) |>
    visLayout(improvedLayout = TRUE)
}, error = function(e) {
  message("Pipeline graph not shown: ", conditionMessage(e))
  message("Install the crew package (e.g. install.packages(\"crew\")) to enable tar_visnetwork().")
})
```

```{r}
#| label: pipeline
#| echo: false
tryCatch({
  setwd(root_dir)
  tar_make(reporter = "verbose_positives",
           garbage_collection = TRUE)
}, error = function(e) {
  cat("Note: Pipeline execution encountered an error:", conditionMessage(e), "\n")
  cat("Some targets may not be available. Check tar_meta() for details.\n")
})
```

```{r}
#| label: crew
#| echo: false
setwd(root_dir)
# Crew workers table: one row per worker (or main process if no crew)
crew_tbl <- tryCatch(
  targets::tar_crew() |> tibble::as_tibble(),
  error = function(e) tibble::tibble(controller = character(), targets = integer(), seconds = numeric())
)
if (nrow(crew_tbl) == 0) {
  # No crew: show main process as single "worker" with summary from progress
  prog_sum <- tryCatch(targets::tar_progress_summary(fields = c("completed", "skipped")), error = function(e) NULL)
  n_done <- NA_integer_
  if (!is.null(prog_sum) && nrow(prog_sum) > 0) {
    n_done <- as.integer(sum(c(prog_sum$completed, prog_sum$skipped), na.rm = TRUE))
  }
  crew_tbl <- tibble::tibble(
    controller = "Main process (no crew)",
    worker = 1L,
    targets = n_done,
    seconds = NA_real_
  )
} else {
  # Expand to one row per worker using controller's worker count
  n_workers <- tryCatch({
    ctrl <- targets::tar_option_get("controller")
    if (!is.null(ctrl) && inherits(ctrl, "crew_class_controller")) as.integer(ctrl$workers) else 1L
  }, error = function(e) 1L)
  crew_tbl <- crew_tbl %>%
    dplyr::slice(rep(1:dplyr::n(), each = n_workers)) %>%
    dplyr::mutate(worker = rep(seq_len(n_workers), length.out = dplyr::n())) %>%
    dplyr::relocate(worker, .after = controller)
}
crew_tbl <- crew_tbl %>%
  dplyr::rename(Controller = controller, Worker = worker, `Targets run` = targets, `Seconds` = seconds)
cat("**Crew workers** (one row per worker):\n\n")
crew_tbl %>% knitr::kable(format = "html", escape = FALSE, table.attr = 'class="table" style="table-layout:fixed; width:100%; max-width:100%"')
```

</details>

<details><summary>Setup unit tests</summary>

```{r}
#| echo: false
tryCatch({
  setwd(root_dir)
  pt <- tar_read(setup_unit_tests)
  pt %>% dplyr::select(test, Passed = passed_html) %>% knitr::kable(format = "html", escape = FALSE, table.attr = 'class="table" style="table-layout:fixed; width:100%; max-width:100%"')
}, error = function(e) {
  cat("Setup unit tests not available. Run the pipeline (tar_make) first.\n")
})
```

</details>

<!-- #################### REPORT #################### -->

```{r}
#| label: data-source-notice
#| echo: false
#| results: asis
# Show notice when pipeline uses anonymised/synthetic data (e.g. from structure.rds).
tryCatch({
  ds <- tar_read(data_source)
  if (isTRUE(ds$use_anonymised)) {
    cat('::: {.callout-note}\n## Data source\nThis report uses **anonymised/synthetic data** (simulated from structure). No real observations are shown.\n:::\n')
  }
}, error = function(e) { invisible(NULL) })
```

```{r}
#| label: doc-labels
#| echo: false
# Document-level variable labels: always use anonymised names (v1..v4).
doc_outcome <- "v1"
doc_covariate <- "v2"
doc_mediator <- "v3"
doc_exposure <- "v4"
doc_log_outcome <- "log(v1)"
tryCatch({
  ds <- tar_read(data_source)
  doc_dataset_name <- if (isTRUE(ds$use_anonymised)) "synthetic data" else "imported data"
}, error = function(e) { doc_dataset_name <- "the dataset" })
```

# Experiment and Data

The analysis uses the **`r doc_dataset_name`** dataset. The outcome is **`r doc_outcome`**, with **`r doc_exposure`** (exposure), **`r doc_mediator`** (mediator), and **`r doc_covariate`** (covariate) as predictors. After cleaning and keeping complete cases on the outcome and covariate, the data were used to model **`r doc_outcome`** as a function of **`r doc_mediator`**, **`r doc_exposure`**, **`r doc_covariate`**, and the **`r doc_mediator`**–**`r doc_exposure`** interaction.

# Methods

A Bayesian regression model was fitted in R (R Core Team 2022) using the **brms** package (Bürkner, 2017) with the **cmdstanr** backend. The model was a **structural equation model** with two equations: (1) **`r doc_mediator`** is regressed on **`r doc_exposure`** (exposure affects mediator); (2) **`r doc_log_outcome`** (outcome log-transformed) is regressed on **`r doc_mediator`**, **`r doc_exposure`**, their **interaction**, and **`r doc_covariate`**, with a Gaussian distribution. So **`r doc_exposure`** affects the outcome both directly and indirectly via **`r doc_mediator`**, and **`r doc_covariate`** is an additional predictor (`r doc_mediator` ~ `r doc_exposure`; `r doc_log_outcome` ~ `r doc_mediator` * `r doc_exposure` + `r doc_covariate`). The model was fitted with 4 chains, 1000 warmup and 2000 iterations per chain. Model validation relied on posterior predictive checks (density overlay) and standard MCMC diagnostics (Rhat and effective sample size) (McElreath, 2020).

<details><summary>Why Bayesian?</summary>
- This framework is particularly useful for the analysis of complex data sets with temporal and spatial (hierarchical) structures and is highly flexible to model the data generating process: <https://doi.org/10.1007/s10144-015-0503-4></details>

## Analysis

::: {.panel-tabset}

## Data

```{r}
#| echo: false
d <- tar_read(data_struct)
str(d)
head(d)
```

## Summary

```{r}
#| echo: false
library(tidyverse)
d <- tar_read(data_struct)
d %>%
  select(where(is.numeric)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  group_by(variable) %>%
  summarise(
    mean = mean(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    range = paste(round(min(value, na.rm = TRUE), 2), "-", round(max(value, na.rm = TRUE), 2)),
    sd = sd(value, na.rm = TRUE),
    .groups = "drop"
  )
```

## Models

```{r}
#| echo: false
tryCatch({
  fit_obj <- tar_read(fit)
  print(fit_obj)
}, error = function(e) {
  cat("Model fit not available. Run the pipeline to generate targets.\n")
})
```

**Stancode**

```{r}
#| echo: false
tryCatch({
  tar_read(stan_code)
}, error = function(e) {
  cat("Stan code not available. Run the pipeline to generate targets.\n")
})
```

## Validation

```{r}
#| echo: false
#| message: false
#| fig-show: hold
#| out-width: 50%
#| fig-width: 4
#| fig-height: 2
tryCatch({
  p <- tar_read(pp_check)
  ggplot2::ggsave(file.path(plot_dir, "pp_check.png"), p, width = 4, height = 2, dpi = 600)
  p
}, error = function(e) {
  cat("Validation plot not available.\n")
})
```

## Analysis unit tests

```{r}
#| echo: false
tryCatch({
  setwd(root_dir)
  pt <- tar_read(unit_tests)
  pt %>% dplyr::select(test, Passed = passed_html) %>% knitr::kable(format = "html", escape = FALSE, table.attr = 'class="table" style="table-layout:fixed; width:100%; max-width:100%"')
}, error = function(e) {
  cat("Analysis unit tests not available. Run the pipeline (tar_make) first.\n")
})
```

:::

```{r}
#| label: export-tables
#| echo: false
#| include: false
# Export final tables to single .xlsx (per project rules: analysis/output/tables/)
setwd(root_dir)
tryCatch({
  library(tidyverse)
  library(writexl)
  d <- tar_read(data_struct)
  summary_tbl <- d %>%
    select(where(is.numeric)) %>%
    pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
    group_by(variable) %>%
    summarise(
      mean = mean(value, na.rm = TRUE),
      median = median(value, na.rm = TRUE),
      range = paste(round(min(value, na.rm = TRUE), 2), "-", round(max(value, na.rm = TRUE), 2)),
      sd = sd(value, na.rm = TRUE),
      .groups = "drop"
    )
  analysis_unit_tests_tbl <- tar_read(unit_tests)
  setup_unit_tests_tbl <- tar_read(setup_unit_tests)
  tables_path <- file.path(tables_dir, "1_targets_report_tables.xlsx")
  writexl::write_xlsx(list(Summary = summary_tbl, Analysis_unit_tests = analysis_unit_tests_tbl, Setup_unit_tests = setup_unit_tests_tbl), tables_path)
}, error = function(e) {
  message("Could not export tables to .xlsx: ", conditionMessage(e))
})
```

### Causal diagram
<details><summary>i</summary>
- A causal diagram (DAG) is a flowchart that maps out the causal relationships between the exposure (X) and outcome (Y), along with additional variables that may affect the X and Y association.

- As such a causal diagram is a graphical representation of the hypothesis and summary of a modeling approach.

- A biologist's guide to model selection and causal inference: <https://doi.org/10.1098/rspb.2020.2815></details>
```{r}
#| echo: false
#| out-width: 50%
#| fig-height: 4.5
#| fig-width: 9
#| fig-align: center
setwd(root_dir)
tryCatch({
  d <- tar_read(dag)
  ggplot2::ggsave(file.path(plot_dir, "causal_dag.png"), d, width = 6, height = 4.5, dpi = 600)
  d
}, error = function(e) {
  cat("DAG not available.\n")
})
```

Figure 1: Causal diagram (DAG): exposure (`r doc_exposure`), mediator (`r doc_mediator`), covariate (`r doc_covariate`), and outcome (`r doc_log_outcome`).

### Raw plots
```{r}
#| echo: false
#| message: false
#| fig-width: 8
#| fig-height: 7
#| out-width: 100%
tryCatch({
  setwd(root_dir)
  p_wind <- tar_read(plot_ozone_wind)
  p_temp <- tar_read(plot_ozone_temp)
  p_solar <- tar_read(plot_ozone_solar_r)
  p_combined <- p_wind + p_temp + p_solar + plot_layout(ncol = 2)
  ggplot2::ggsave(file.path(plot_dir, "v1_vs_v3_raw.png"), p_wind, width = 4, height = 3.5, dpi = 600)
  ggplot2::ggsave(file.path(plot_dir, "v1_vs_v4_raw.png"), p_temp, width = 4, height = 3.5, dpi = 600)
  ggplot2::ggsave(file.path(plot_dir, "v1_vs_v2_raw.png"), p_solar, width = 4, height = 3.5, dpi = 600)
  ggplot2::ggsave(file.path(plot_dir, "raw_plots_v1_v2_v3_v4.png"), p_combined, width = 8, height = 7, dpi = 600)
  p_combined
}, error = function(e) {
  cat("Raw plots not available. Run the pipeline (tar_make) first.\n")
})
```
Figure 2: Observed `r doc_outcome` (points) and posterior predictions from the Gaussian model for `r doc_log_outcome`, back-transformed to `r doc_outcome` scale. Left: `r doc_outcome` vs `r doc_mediator`. Centre: `r doc_outcome` vs `r doc_exposure`. Right: `r doc_outcome` vs `r doc_covariate`. Shaded bands show the 95% credible interval; the solid line is the posterior median.

### Effect plots
```{r}
#| echo: false
#| message: false
#| fig-show: hold
#| fig-width: 8
#| fig-height: 4
#| out-width: 100%
#| classes: "fig-squeeze"
tryCatch({
  setwd(root_dir)
  p_eff_temp <- tar_read(eff_wind_by_temp)
  p_eff_solar <- tar_read(eff_wind_by_solar_r)
  ggplot2::ggsave(file.path(plot_dir, "effect_v3_by_v4.png"), p_eff_temp, width = 8, height = 4, dpi = 600)
  ggplot2::ggsave(file.path(plot_dir, "effect_v3_by_v2.png"), p_eff_solar, width = 8, height = 4, dpi = 600)
  p_eff_temp
}, error = function(e) {
  cat("Effect plot not available.\n")
})
```
Figure 3: Interaction effect of `r doc_mediator` on `r doc_outcome` at low vs high `r doc_exposure` (25th and 75th percentiles). Grey shading is the 95% credible interval; the black line is the posterior median. Legend indicates `r doc_exposure` level.

```{r}
#| echo: false
#| message: false
#| fig-show: hold
#| fig-width: 8
#| fig-height: 4
#| out-width: 100%
#| classes: "fig-squeeze"
tryCatch({
  setwd(root_dir)
  p_eff_solar <- tar_read(eff_wind_by_solar_r)
  p_eff_solar
}, error = function(e) {
  cat("Effect plot (mediator by covariate) not available.\n")
})
```
Figure 4: Effect of `r doc_mediator` on `r doc_outcome` at low vs high `r doc_covariate` (25th and 75th percentiles). Grey shading is the 95% credible interval; the black line is the posterior median. Legend indicates `r doc_covariate` level.

## Results

The causal diagram (Figure 1) encodes the assumption that `r doc_mediator`, `r doc_exposure`, and `r doc_covariate` influence `r doc_outcome`. The raw plots (Figure 2, 2-column layout) show that observed `r doc_outcome` tends to decrease with higher `r doc_mediator` and to increase with higher `r doc_exposure` and `r doc_covariate`, with the model capturing these patterns and uncertainty. The interaction effect plot (Figure 3) shows how the effect of `r doc_mediator` on `r doc_outcome` differs at low vs high `r doc_exposure`; Figure 4 shows the effect of `r doc_mediator` at low vs high `r doc_covariate`. Non-parallel curves indicate an interaction. In this Bayesian framework, a predictor is typically considered to have a meaningful effect when its 95% credible interval excludes zero.

# References

Bürkner, P.C., (2017). brms: An R Package for Bayesian Multilevel Models Using Stan. Journal of Statistical Software, 80(1), 1-28. <https://econpapers.repec.org/article/jssjstsof/v_3a080_3ai01.htm>

McElreath, R. (2020). Statistical rethinking: A Bayesian course with examples in R and Stan. Chapman and Hall/CRC.

R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>

<details><summary>Report unit tests</summary>

```{r}
#| echo: false
tryCatch({
  setwd(root_dir)
  pt <- tar_read(report_unit_tests)
  pt %>% dplyr::select(test, Passed = passed_html) %>%
    knitr::kable(format = "html", escape = FALSE, table.attr = 'class="table" style="table-layout:fixed; width:100%; max-width:100%"')
}, error = function(e) {
  cat("Report unit tests not available. Run the pipeline (tar_make) first.\n")
})
```

</details>

<details><summary>Session Information</summary>
```{r}
#| label: session-info
# Template version 2.6
utils::sessionInfo()
```
</details>
